### 2016-09-27 revision/solo-ftrl.tex
Improvements

In this section, we introduces our first scale-free algorithm, based on \textsc{FTRL}.

The closest algorithm to a scale-free \textsc{FTRL} in the existing literature
is the \textsc{AdaGrad FTRL} algorithm~\cite{Duchi-Hazan-Singer-2011}. It uses a
regularizer on each coordinate of the form
\begin{equation*}
R_t(w) = R(w) \left(\delta + \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \right).
\end{equation*}
This kind of regularizer would yield a scale-free algorithm \emph{only} for
$\delta=0$.  Unfortunately, the regret bound in~\cite{Duchi-Hazan-Singer-2011}
becomes vacuous for such setting in the unbounded case. In fact, it requires
$\delta$ to be greater than $\|\ell_t\|_*$ for all time steps $t$, requiring
knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
other words, despite of its name, \textsc{AdaGrad} is not fully adaptive to the
norm of the gradient vectors. Identical considerations hold for the
\textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: the
scale-free setting of the learning rate is valid only in the bounded case.

@@

When $K$ is unbounded, we pay a penalty that scales as $\max_{t \le T} \|\ell_t\|_* \sqrt{T}$, that has the same magnitude of the first term in the bound.
On the other hand, when $K$ is bounded, the second term is a constant and we can choose the optimal multiple of the regularizer.  We
choose $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex function and
optimize $\lambda$.  The result of the optimization is
Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}.
%; the proof is in~\ref{section:solo-ftrl-proofs}.
%It is similar to Corollary~\ref{corollary:ada-ftrl-regret-bound} for \textsc{AdaFTRL}. The scaling however is different in the two corollaries.  In Corollary~\ref{corollary:ada-ftrl-regret-bound}, $\lambda \sim 1/(\sup_{v \in K} f(v))$ while in Corollary~\ref{corollary:regret-solo-ftrl-bounded-set} we have $\lambda \sim 1/\sqrt{\sup_{v \in K} f(v)}$.

@@

The proof of Theorem~\ref{theorem:regret-solo-ftrl} relies on an inequality
(Lemma~\ref{lemma:useful}).  Related and weaker inequalities were proved
in~\cite{Auer-Cesa-Bianchi-Gentile-2002} and~\cite{Jaksch-Ortner-Auer-2010}.
The main property of this inequality is that on the right-hand side $C$ does
\emph{not} multiply the $\sqrt{\sum_{t=1}^T a_t^2}$ term.  We will also use the
well-known technical Lemma~\ref{lemma:sum-of-square-roots-inverses}.

@@

\begin{lemma}[{{\cite[Lemma~3.5]{Auer-Cesa-Bianchi-Gentile-2002}}}]
\label{lemma:sum-of-square-roots-inverses}
Let $a_1, a_2, \dots, a_T$ be non-negative real numbers. If $a_1 > 0$ then,
$$
\sum_{t=1}^T \frac{a_t}{\sqrt{\sum_{i=1}^t a_i}} \le 2 \sqrt{\sum_{t=1}^T a_t} \; .
$$
\end{lemma}
%
For completeness, a proof of Lemma~\ref{lemma:sum-of-square-roots-inverses} is in
in~\ref{section:solo-ftrl-proofs}.

>>>>

In this section, we introduce our first scale-free algorithm; it will be based
on \textsc{FTRL}. The closest algorithm to a scale-free \textsc{FTRL} in the
existing literature is the \textsc{AdaGrad FTRL}
algorithm~\cite{Duchi-Hazan-Singer-2011}. It uses a regularizer on each
coordinate of the form
\begin{equation*}
R_t(w) = R(w) \left(\delta + \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \right).
\end{equation*}
This kind of regularizer would yield a scale-free algorithm \emph{only} for
$\delta=0$.  Unfortunately, the regret bound in~\cite{Duchi-Hazan-Singer-2011}
becomes vacuous for such setting in the unbounded case. In fact, it requires
$\delta$ to be greater than $\|\ell_t\|_*$ for all time steps $t$, requiring
knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
other words, despite of its name, \textsc{AdaGrad FTRL} is not fully adaptive to
the norm of the gradient vectors. Identical considerations hold for the
\textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: the
scale-free setting of the learning rate is valid only in the bounded case.

@@

When $K$ is unbounded, we pay a penalty that scales as $\max_{t \le T}
\|\ell_t\|_* \sqrt{T}$, that has the same magnitude of the first term in the
bound. On the other hand, when $K$ is bounded, the second term is a constant and
we can choose the optimal multiple of the regularizer.  We choose $R(w) =
\lambda f(w)$ where $f$ is a $1$-strongly convex function and optimize
$\lambda$.  The result of the optimization is
Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}.

@@

The proof of Theorem~\ref{theorem:regret-solo-ftrl} relies on an inequality
(Lemma~\ref{lemma:useful}).  Related and weaker inequalities
(Lemma~\ref{lemma:sum-of-square-roots-inverses}) were proved
in~\cite{Auer-Cesa-Bianchi-Gentile-2002} and~\cite{Jaksch-Ortner-Auer-2010}. The
main property of this inequality is that on the right-hand side $C$ does
\emph{not} multiply the $\sqrt{\sum_{t=1}^T a_t^2}$ term.

@@

\begin{lemma}[{{\cite[Lemma~3.5]{Auer-Cesa-Bianchi-Gentile-2002}}}]
\label{lemma:sum-of-square-roots-inverses}
Let $a_1, a_2, \dots, a_T$ be non-negative real numbers. If $a_1 > 0$ then,
$$
\sum_{t=1}^T \frac{a_t}{\sqrt{\sum_{i=1}^t a_i}} \le 2 \sqrt{\sum_{t=1}^T a_t} \; .
$$
\end{lemma}
%
For completeness, a proof of Lemma~\ref{lemma:sum-of-square-roots-inverses} is
in~\ref{section:solo-ftrl-proofs}.
